Channel Shaping Using Reconfigurable Intelligent Surfaces: From Diagonal to Beyond

### 2024-01-09 main.tex
:pencil: Add contributions

	In conventional (a.k.a. diagonal) \gls{ris}, each scattering element is tuned by a dedicated impedance and functions as an \emph{individual} phase shifter \cite{Wu2020}.
	% , accounting for a diagonal phase shift model with unit magnitude.
	The idea is extended to \gls{bd} \gls{ris} \cite{Shen2020a,Li2023b} featuring in-group connections between elements.
	This allows \emph{cooperative} reflection --- wave impinging on one element can propagate within the circuit and depart partially from all elements.
	\gls{bd} \gls{ris} can thus control both amplitude and phase of the reflected wave, generalizing the scattering matrix from diagonal with unit-magnitude entries to block diagonal with  unitary blocks.
	Its benefit has been recently shown in receive power maximization \cite{Nerini2023,Santamaria2023,Fang2023,Nerini2023a}, transmit power minimization \cite{Zhou2023}, and rate maximization \cite{Zhou2023,Nerini2023a,Li2023d,Li2023c,Bartoli2023}.
	The channel estimation \cite{Li2023e} and mutual coupling \cite{Li2023f} issues have also been investigated.
	Therefore, \gls{bd} \gls{ris} is envisioned as the next generation channel shaper with stronger signal processing flexibility \cite{Li2023g}.

	% Attempts to characterize the channel shaping capability can be classified into \emph{singular value centric} and \emph{power centric} methods.
	Metrics of channel shaping can be classified into \emph{singular value centric} and \emph{power centric}.
	For diagonal \gls{ris}, the former has been studied in terms of minimum singular value \cite{ElMossallamy2021}, effective rank \cite{ElMossallamy2021,Meng2023}, condition number \cite{Zheng2022,Huang2023} in \gls{pc}, and degree of freedom \cite{Bafghi2022,Zheng2023,Chae2023} in \gls{ic}.
	The latter has been studied in terms of channel power \cite{Wu2019} in \gls{pc} and leakage interference \cite{Santamaria2023a} in \gls{ic}.
	When it comes to \gls{bd} \gls{ris}, the only metric analyzed is the channel power \cite{Nerini2023,Santamaria2023} where the \gls{mimo} case is still unexplored.\footnote{Single-stream \gls{mimo} with given precoder and combiner was considered in \cite{Nerini2023}. It is equivalent \gls{siso} channel shaping.}
	Although insightful, the answers are far from complete and the optimization techniques are limited to very specific scenarios.
	In this paper, we aim to quantify the channel shaping capability of \gls{bd} \gls{ris} through theoretical analysis and numerical evaluation.
	The contributions are summarized below.

	% we aim to answer this question by a simple yet insightful approach: \emph{from diagonal to beyond}.
	% The answer depends heavily on the phase shift model of \gls{ris}.

	% \gls{ris} also offers a new degree of freedom for \emph{channel shaping}, i.e., to reshape the wireless channel to a desired form.
\end{section}

@@

\begin{section}{\glsentryshort{mimo}-\glsentryshort{pc}}
	\begin{subsection}{Channel Power Maximization}
		Consider a \gls{bd} \gls{ris} with $N^\mathrm{S}$ elements, which is divided into $G$ groups of equal $L$ elements.
		\begin{maxi!}
			{\scriptstyle{\mathbf{\Theta}}}{\left\lVert \mathbf{H}^\mathrm{D} + \sum_g\nolimits \mathbf{H}_g^\mathrm{B} \mathbf{\Theta}_g \mathbf{H}_g^\mathrm{F} \right\rVert _\mathrm{F}^2}{\label{op:pc_power}}{}
			\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I}, \quad \forall g \in \mathcal{G} \triangleq \{1,\ldots,G\}.}{}{}
		\end{maxi!}
		For \emph{symmetric} BD-RIS, the problem has been solved in
		\begin{itemize}
			\item Matteo's paper \cite{Nerini2023}: SISO and equivalent\footnote{Single-stream MIMO with given precoder and combiner.};
			\item Ignacio's paper \cite{Santamaria2023}: SISO and directless MISO/SIMO.
		\end{itemize}

@@

		Finally, we characterize the \emph{Pareto front} of channel singular values via optimization approach.
		% \begin{maxi!}
		% 	{\scriptstyle{\mathbf{\Theta}}}{J_1 = \sum_n \rho_n \sigma_n(\mathbf{H})}{\label{op:pc_singular}}{\label{ob:pc_singular}}
		% 	\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I}, \quad \forall g,}{}{}
		% \end{maxi!}

		\begin{figure}[!t]
			\centering
			\resizebox{0.65\columnwidth}{!}{
				\input{assets/simulation/pc_singular_pareto.tex}
			}
			\caption{Singular value Pareto front. $(N^\mathrm{T}, N^\mathrm{S}, N^\mathrm{R}) = (4, 64, 2)$, $(\Lambda^\mathrm{D}, \Lambda^\mathrm{F}, \Lambda^\mathrm{B}) = (0, -17.5, -17.5) \unit{dB}$.}
			\label{sm:pc_singular_pareto}
		\end{figure}

		The Pareto front and evolving trend of channel singular values are shown in Fig.~\ref{sm:pc_singular_pareto} and \ref{sm:pc_singular_bound}.
		Clearly, \gls{bd} \gls{ris} with a larger group size can redistribute the channel singular values to a wider range.
	\end{subsection}
\end{section}

>>>>

	In conventional (a.k.a. diagonal) \gls{ris}, each scattering element is tuned by a dedicated impedance and functions as an \emph{individual} phase shifter \cite{Wu2020}.
	% , accounting for a diagonal phase shift model with unit magnitude.
	The idea is extended to \gls{bd} \gls{ris} \cite{Shen2020a,Li2023b} featuring in-group connections between elements.
	This allows \emph{cooperative} reflection --- wave impinging on one element can propagate within the circuit and depart partially from all elements.
	\gls{bd} \gls{ris} can thus control both amplitude and phase of the reflected wave, generalizing the scattering matrix from diagonal with unit-magnitude entries to block diagonal with  unitary blocks.
	Its benefit has been recently shown in receive power maximization \cite{Nerini2023,Santamaria2023,Fang2023,Nerini2023a}, transmit power minimization \cite{Zhou2023}, and rate maximization \cite{Zhou2023,Nerini2023a,Li2023d,Bartoli2023,Li2023c}.
	The channel estimation \cite{Li2023e} and mutual coupling \cite{Li2023f} issues have also been investigated.
	Therefore, \gls{bd} \gls{ris} is envisioned as the next generation channel shaper with stronger signal processing flexibility \cite{Li2023g}.

	% Attempts to characterize the channel shaping capability can be classified into \emph{singular value centric} and \emph{power centric} methods.
	Channel shaping metrics can be classified into \emph{singular value centric} and \emph{power centric}.
	For diagonal \gls{ris}, the former has been studied in terms of minimum singular value \cite{ElMossallamy2021}, effective rank \cite{ElMossallamy2021,Meng2023}, condition number \cite{Zheng2022,Huang2023} in \gls{pc}, and degree of freedom \cite{Bafghi2022,Zheng2023,Chae2023} in \gls{ic}.
	The latter has been studied in terms of channel power \cite{Wu2019} in \gls{pc} and leakage interference \cite{Santamaria2023a} in \gls{ic}.
	When it comes to \gls{bd} \gls{ris}, the only metric analyzed is the channel power \cite{Nerini2023,Santamaria2023} while the \gls{mimo} case is still unexplored.\footnote{Single-stream \gls{mimo} with given precoder and combiner was considered in \cite{Nerini2023}. In terms of channel shaping, it is equivalent to \gls{siso}.}
	Although insightful, those attempts are far from complete and the methods are limited to specific scenarios.
	% This paper aims to quantify the channel shaping capability of asymmetric \gls{bd} \gls{ris} through theoretical analysis and numerical evaluation.
	This paper aims for a comprehensive answer to the channel shaping question through theoretical analysis and numerical optimization.
	% quantify the channel shaping capability of asymmetric \gls{bd} \gls{ris} through theoretical analysis and numerical evaluation.
	The contributions are summarized below.

	First, we examine the capability of \gls{bd} \gls{ris} to redistribute the channel singular values in \gls{mimo} \gls{pc}.
	The Pareto frontiers are characterized by optimizing the \emph{weighted sum of singular values}, where the weights can be positive, zero, or negative.
	It generalizes all singular value metrics and provides a powerful design framework.
	We then discuss analytical singular value bounds under rank-deficient and no-direct scenarios.

	% and demonstrate the channel shaping advantage of \gls{bd} \gls{ris} over diagonal \gls{ris}.

	Second, we tailor a \gls{rcg} algorithm for asymmetric\footnote{Although the constraint of symmetric scattering parameters is widely respected in the literature \cite{Shen2020a,Nerini2023,Santamaria2023,Fang2023,Nerini2023a,Zhou2023,Li2023d,Bartoli2023}, it can be relaxed when the \gls{bd} \gls{ris} involves asymmetric passive components \cite{Ahn2006}, as previously assumed in \cite{Li2023b,Li2023c}. Symmetry can be enforced by projection if necessary.} \gls{bd} \gls{ris} optimization.
	Specifically, block-wise update is performed along the geodesics\footnote{A geodesic refers to the shortest path between two points in a Riemannian manifold.} of the Stiefel manifold and concisely evaluated as matrix exponential.
	The proposed method not only features lower complexity and faster convergence than universal manifold optimization \cite{Absil2009,Pan2022d}, but also applies to any \gls{bd} \gls{ris} problem where gradient exists.
	% It is used to solve the Pareto singular problem, as well as rate maximization in \gls{mimo} \gls{pc} and \gls{wsr} maximization in \gls{mimo} \gls{ic}.
	% It effectively solves the Pareto singular problem.

	% , as well as rate maximization in \gls{mimo} \gls{pc} and \gls{wsr} maximization in \gls{mimo} \gls{ic}.

	% It helps to solve the aforementioned Pareto singular value problem.
	% We exploit it to solve the Pareto singular value problem and

	% It is exploited for the Pareto singular value problem, as well as rate maximization in \gls{mimo} \gls{pc} and \gls{wsr} maximization in \gls{mimo} \gls{ic}.
	% It is used to solve the Pareto singular value problem,

	Third, we consider power centric design
	% Second, we consider power centric design
	% exploit \gls{bd} \gls{ris} to power centric
	% maximize channel power in \gls{mimo} \gls{pc} and minimize leakage interference in \gls{mimo} \gls{ic}.

\end{section}

@@

\begin{section}{\glsentryshort{mimo}-\glsentryshort{pc}}

		Finally, we characterize the \emph{Pareto frontier} of channel singular values via optimization approach.
		% \begin{maxi!}
		% 	{\scriptstyle{\mathbf{\Theta}}}{J_1 = \sum_n \rho_n \sigma_n(\mathbf{H})}{\label{op:pc_singular}}{\label{ob:pc_singular}}
		% 	\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I}, \quad \forall g,}{}{}
		% \end{maxi!}

		\begin{figure}[!t]
			\centering
			\resizebox{0.65\columnwidth}{!}{
				\input{assets/simulation/pc_singular_pareto.tex}
			}
			\caption{Singular value Pareto frontier. $(N^\mathrm{T}, N^\mathrm{S}, N^\mathrm{R}) = (4, 64, 2)$, $(\Lambda^\mathrm{D}, \Lambda^\mathrm{F}, \Lambda^\mathrm{B}) = (0, -17.5, -17.5) \unit{dB}$.}
			\label{sm:pc_singular_pareto}
		\end{figure}

		The Pareto frontier and evolving trend of channel singular values are shown in Fig.~\ref{sm:pc_singular_pareto} and \ref{sm:pc_singular_bound}.
		Clearly, \gls{bd} \gls{ris} with a larger group size can redistribute the channel singular values to a wider range.
	\end{subsection}
	\begin{subsection}{Channel Power Maximization}
		Consider a \gls{bd} \gls{ris} with $N^\mathrm{S}$ elements, which is divided into $G$ groups of equal $L$ elements.
		\begin{maxi!}
			{\scriptstyle{\mathbf{\Theta}}}{\left\lVert \mathbf{H}^\mathrm{D} + \sum_g\nolimits \mathbf{H}_g^\mathrm{B} \mathbf{\Theta}_g \mathbf{H}_g^\mathrm{F} \right\rVert _\mathrm{F}^2}{\label{op:pc_power}}{}
			\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I}, \quad \forall g \in \mathcal{G} \triangleq \{1,\ldots,G\}.}{}{}
		\end{maxi!}
		For \emph{symmetric} BD-RIS, the problem has been solved in
		\begin{itemize}
			\item Matteo's paper \cite{Nerini2023}: SISO and equivalent\footnote{Single-stream MIMO with given precoder and combiner.};
			\item Ignacio's paper \cite{Santamaria2023}: SISO and directless MISO/SIMO.
		\end{itemize}

@@

\end{section}
